---
title: "Reproducible Research: Peer Assessment 1"
author : "Phillip Hardwick"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
reprorepo <- read.csv("activity.csv")
uni_d <- unique(reprorepo$date)

dsums_dt <- data.frame(uni_d)
for(i in 1:length(dsums_dt$uni_d)){
  dsums_dt$sums[i] <- sum(reprorepo$steps[which(reprorepo$date==dsums_dt$uni_d[i])])
  dsums_dt$median[i] <- median(reprorepo$steps[which(reprorepo$date==dsums_dt$uni_d[i])])
  dsums_dt$mean[i] <- mean(reprorepo$steps[which(reprorepo$date==dsums_dt$uni_d[i])])
}
```


## histogram of the total number of steps taken each day

```{r}
hist(dsums_dt$sums)
```


## Median, Mean, and Sum Values per Day
```{r}
dsums_dt
```

```{r}
repro_interval_dt <- data.frame(int = unique(reprorepo$interval))
for(i in 1: length(repro_interval_dt$int)){
  repro_interval_dt$int_mean[i] <- mean(reprorepo$steps[which(reprorepo$interval==repro_interval_dt$int[i])],na.rm = TRUE)
}
```


## What is mean total number of steps taken per day?

## Average number of steps taken averaged across all days

```{r}
plot(repro_interval_dt$int_mean~repro_interval_dt$int,type="l",ylab="Average Number of Steps",xlab="Time Interval", asp=4,col.lab = "blue",cex=1)

```


## Interval with highest number of steps
```{r}
repro_interval_dt$int[which(repro_interval_dt$int_mean == max(repro_interval_dt$int_mean))]
```


## What is the average daily activity pattern?



## Imputing missing values

## Number of NA values in dataset
```{r}
  sum(is.na(reprorepo$steps))
```

## Missing values filled in by using mean data per interval, new copy of dataset created, mean and median calculated
```{r}
reprorepo_copy <- reprorepo
reprorepo_copy$steps <- with(reprorepo, ave(steps, interval,FUN = function(x) replace(x, is.na(x), mean(x, na.rm = TRUE))))
newsums_dt <- data.frame(uni_d)
for(i in 1:length(newsums_dt$uni_d)){
  newsums_dt$sums[i] <- sum(reprorepo_copy$steps[which(reprorepo_copy$date==newsums_dt$uni_d[i])])
  newsums_dt$median[i] <- median(reprorepo_copy$steps[which(reprorepo_copy$date==newsums_dt$uni_d[i])])
  newsums_dt$mean[i] <- mean(reprorepo_copy$steps[which(reprorepo_copy$date==newsums_dt$uni_d[i])])
}
```

## Mean and median values 

```{r}
  newsums_dt
```

## Histogram of data with NA data replaced
```{r}
  hist(newsums_dt$sums)
```

## Are there differences in activity patterns between weekdays and weekends?

## Derive day of week from Date and assign factor of either weekend or weekday
```{r}
  reprorepo_copy$date <- as.Date(reprorepo_copy$date,"%Y-%m-%d")
weekend_days <- c("Saturday","Sunday")
reprorepo_copy$part_of_week <- factor((weekdays(reprorepo_copy$date) %in% weekend_days), levels=c(FALSE, TRUE), labels=c('weekday', 'weekend'))
repro_nona_weekend <- data.frame(int = unique(reprorepo_copy$interval))   
repro_nona_weekday <- data.frame(int = unique(reprorepo_copy$interval))
for(i in 1: length(repro_nona_weekend$int)){
  repro_nona_weekend$int_mean[i] <-mean(reprorepo_copy$steps[which(reprorepo_copy$part_of_week == 'weekend' & reprorepo_copy$interval == repro_nona_weekend$int[i])])

}
for(i in 1: length(repro_nona_weekday$int)){
  repro_nona_weekday$int_mean[i] <-mean(reprorepo_copy$steps[which(reprorepo_copy$part_of_week == 'weekday' & reprorepo_copy$interval == repro_nona_weekday$int[i])])
  
}
```

## Plot comparing weekend and weekday activity
```{r}
par(mfrow = c(2, 1))
par(cex = .4)
par(col.lab = "red")
plot(repro_nona_weekday$int_mean~repro_nona_weekday$int,type="l",main = "weekday",ylab="Average Number of Steps",xlab = "",asp = 4)
plot(repro_nona_weekend$int_mean~repro_nona_weekend$int,type="l",main = "weekend",ylab="Average Number of Steps",xlab="Time Interval", asp = 4)
```